Draft

### 2017-08-27 Draft/Draft.tex
updated figures

\title{Literature Review}

@@

The ability to forecast exchange rates is of great importance to policy
makers, businesses and speculators. Policy makers are interested in
forecasting exchange rates to assess how its movement will impact
inflation, and how monetary policy will need to be adjusted in the
future to keep inflation within the inflation target (Fahimifard et al.
\protect\hyperlink{ref-fahimifard2009}{2009}). Businesses need to assess
how future movements in exchange rates will affect the cost of importing
goods and the demand for goods exported, while speculators aim to
identify opportunities in price movements which can be exploited to make
a profit.

Producing accurate forecasts of exchange rates has proven to be a
difficult task due to the noise of the time series (Abu-Mostafa and
Atiya \protect\hyperlink{ref-abu1996}{1996}). Some economists argue that
the exchange rate market is very efficient and future movements in price
are unpredictable (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}). This belief in the efficient
market hypothesis has led many researchers to model exchange rates using
a Random Walk (RW) model (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}). Although the RW model has
produced superior forecasts, some researchers believe that future prices
depend on present and past events, and other techniques such as the
autoregressive integrated moving average (ARIMA) model has been applied
to forecast exchange rates (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}).

The ARIMA model is one of the most commonly used methods in financial
forecasting (Fahimifard et al.
\protect\hyperlink{ref-fahimifard2009}{2009}). It is known for its
attractive quality of being a rich class of processes, making it
possible to find a model that adequately fits the data (Fahimifard et
al. \protect\hyperlink{ref-fahimifard2009}{2009}). Some research has
shown that the nature of exchange rate data is non-linear, and this
non-linearity may not be captured adequately by the ARIMA model (Zhang
and Hu \protect\hyperlink{ref-zhang1998}{1998}). Linear parametric
forecasting techniques such as the generalised autoregressive
conditional heteroskedasticity (GARCH) model has been proposed as a
solution to this problem (Zhang and Hu
\protect\hyperlink{ref-zhang1998}{1998}). Although the GARCH model takes
care of the non-linearity in the data, the pre-specification of the
model may prevent it from capturing all the different non-linearities
which may be present in the data.

With the evolution of computational power, other non-linear
non-parametric techniques such as Artificial Neural Networks (ANN),
Support Vector Regression (SVR), Fuzzy Logistic Systems and hybrid
versions of the ARIMA model have been proposed. The ARIMA model has been
used as the standard benchmark model against which these other methods
have been compared. Although they have produced significantly better
results than the ARIMA model, building these models is a complex task.
The ARIMA model is more tractable, less computationally expensive, and
has been used as the building blocks for more advanced models such as
the autoregressive fractionally integrated (ARFIMA), Fuzzy ARIMA, and
seasonal ARIMA (SARIMA) model which have been used for exchange rate
forecasting.

Facebook recently open sourced their forecasting model Prophet. Prophet
is said to be preferable over the ARIMA model as it has the ability to
forecast at scale, is easily interpretable and has parameters that can
be intuitively adjusted (Taylor and Letham
\protect\hyperlink{ref-taylor2017}{2017}). This paper aims to compare
the forecasts produced by the traditional ARIMA model and Prophet
through an evaluation of Bitcoin/ZAR, using a Mincer-Zarnowitz approach
of measuring forecast accuracy. Section 2 will consider the origins of
the ARIMA model and how it has performed and evolved over time by
exploring empirical studies where it has been used for forecasting
exchange rates. Section 3 will also consider other techniques that have
been used to forecast exchange rates and how they have been compared to
the ARIMA model. Section 4 will look at hybrid models that were inspired
by the ARIMA model, while section 5 looks at forecasting with Prophet.
Finally, section 6 will look at the Mincer-Zarnowitz approach to
evaluating forecasts.

\section{Forecasting Exchange Rates with the ARIMA
Model}\label{forecasting-exchange-rates-with-the-arima-model}

\subsection{Birth of ARIMA}\label{birth-of-arima}

@@

\[ 
    y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\).

@@

\[ 
    (1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\).

\subsection{Empirical Applications}\label{empirical-applications}

ARIMA models have been commonly used in financial forecasting and are
popular for observing stock prices and exchange rates due to its power
and statistical properties (C.-S. Lin, Chiu, and Lin
\protect\hyperlink{ref-lin2012}{2012}). They have been used as a
benchmark to compare new forecasting techniques which have emerged over
time. Kamruzzaman and Sarker
(\protect\hyperlink{ref-kamruzzaman2003}{2003}) applied an ARIMA model
to forecast the exchange rate between the Australian dollar and six
other currencies. Its results were used to evaluate the performance of
three Neural Network models. The Normalised Mean Square Error (NMSE) and
Mean Absolute Error (MAE) were used to measure the forecast errors while
the Directional Symmetry (DS), Correct Up trend (CU) and Correct Down
trend (CD) were used to measure the accuracy of the direction of the
forecasts. Similarly, C.-S. Lin, Chiu, and Lin
(\protect\hyperlink{ref-lin2012}{2012}) employed an ARIMA model to
forecast the Taiwan New Dollar exchange rate. The aim of applying the
ARIMA model was to analyse the ability of a proposed hybrid model
created from the Squares Support Vector Regression (LSSVR) and Empirical
Mode Decomposition (EMD) model. C.-S. Lin, Chiu, and Lin
(\protect\hyperlink{ref-lin2012}{2012}) used the same indicators as
Kamruzzaman and Sarker (\protect\hyperlink{ref-kamruzzaman2003}{2003})
to evaluate the accuracy of the direction of the forecasts. To measure
the forecast errors produced they applied the mean absolute percentage
error (MAPE), root‐mean‐square error (RMSE) and mean absolute difference
(MAD). Khashei, Bijari, and Ardali
(\protect\hyperlink{ref-khashei2012}{2012}) also investigated the
ability to forecast exchange rates by considering the British pound and
US dollar, over a short and long forecast horizon. They compared the
predictive capability of their proposed hybrid model consisting of an
ARIMA and a Probabilistic Neural Network (PNN), to the traditional ARIMA
model, and justified the use of the ARIMA as a benchmark, by claiming
that it is the most important linear model. The ARIMA model has also
yielded satisfactory results when predicting exchange rates. Nwankwo
(\protect\hyperlink{ref-nwankwo2014}{2014}) forecasted the Nairo/dollar
rate and diagnostic testing revealed that the ARIMA(1,0,0) model was the
best fit for the data based on the AIC.

The ARIMA model is attractive as it is tractable and produces good
forecasts in short time periods where more than 100 observations are
used (Tseng et al. \protect\hyperlink{ref-tseng2001}{2001}). Although
the ARIMA model has the advantage of ease of implementation and
flexibility, it fails to capture the non-linearity and volatility
present in exchange rate data.

\subsection{Evolution of ARIMA}\label{evolution-of-arima}

Over time, the ARIMA model has evolved to cater for a wider variety of
data that it can be applied to, and to compensate for some of its
shortcomings. The most popular versions of the ARIMA model that has been
implemented in exchange rate forecasting is the Seasonal ARIMA (SARIMA)
and Fractional ARIMA model (FARIMA).

The SARIMA model was introduced by Box and Jenkins as a generalisation
of the ARIMA model to capture the periodic behaviour of data. SARIMA
models extend the ARIMA class by including a term for seasonal
differencing. Exchange rate data that has been observed have shown
seasonality, and have been attempted to be fitted by SARIMA models
(Etuk, Wokoma, and Moffat \protect\hyperlink{ref-etuk2013}{2013}). Etuk,
Wokoma, and Moffat (\protect\hyperlink{ref-etuk2013}{2013}) modelled the
Naira/CFA Franc exchange rate which exhibited monthly seasonality using
an additive SARIMA model, to demonstrate that it can be a useful fit to
the data. Their results showed that the SARIMA model adequately
described the variation in the exchange rate series. This agrees with
the results of a study by Etuk and others
(\protect\hyperlink{ref-etuk2012}{2012}) which fitted a multiplicative
SARIMA model to the Naira/Dollar exchange rates.

The ARFIMA model is another generalisation of the ARIMA model in that
the degree required to make the data stationary can assume any real
value, and is no longer restricted to the integer domain. While the
ARIMA model displays exponentially decaying autocorrelation between
observations the further apart they are in time, the ARFIMA model has
the ability to capture the dependence between observations that are
widely spread apart in time (Cheung
\protect\hyperlink{ref-cheung1993}{1993}). This makes the ARFIMA model
more parsimonious since it can capture long memory in data as well as
short term dynamics (Cheung \protect\hyperlink{ref-cheung1993}{1993}).
Cheung (\protect\hyperlink{ref-cheung1993}{1993}) fitted the ARFIMA
model to examine five exchange rates, and found that there was strong
evidence of long memory in the exchange rate time series. Although the
ARFIMA model fitted the data well, it failed to beat the random walk in
out-of sample forecasts.

GARCH models were later developed due to the failure of ARIMA models to
capture the volatility in financial markets (Anastasakis and Mort
\protect\hyperlink{ref-anastasakis2009}{2009}). They have been widely
used to forecast the volatility of stock returns, interest rates and
exchange rates and are useful for portfolio allocation decisions as well
as option pricing (Cermak \protect\hyperlink{ref-cermak2017}{2017}).
Cermak (\protect\hyperlink{ref-cermak2017}{2017}) employed a GARCH(1,1)
model to analyse the volatility of Bitcoin in countries in which Bitcoin
is commonly traded, while Hsieh (1989) applied a GARCH model to
investigate five exchange rates. Hsieh's (1989) results showed that
although the GARCH model outperformed the random walk, some non-linear
information still remained in the residuals. In a similar study,
Fahimifard et al. (\protect\hyperlink{ref-fahimifard2009}{2009})
illustrated that the GARCH model outperformed the ARIMA, while Neural
Networks outperformed both the GARCH and ARIMA model.

\section{Other techniques used to forecast exchange
rates}\label{other-techniques-used-to-forecast-exchange-rates}

Over time, financial forecasting methods have moved away from linear
models like ARIMA and GARCH, to soft computing techniques. These more
complex techniques are non-linear and can fit complex time series more
easily (Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}).
Unlike regression analysis and the ARIMA model, soft computing
techniques do not impose structural assumptions on the model apriori
(Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}). Some of
the most commonly used artificial intelligence methods used to forecast
exchange rate data are Neural Networks and the Fuzzy Logistic Systems.

\subsection{Neural Networks}\label{neural-networks}

Artificial Neural Networks are commonly used in financial forecasting
and are more advantageous than other non-linear forecasting methods.
They can approximate any continuous function, are data driven, and can
adapt to non-stationary environments (Khashei and Bijari
\protect\hyperlink{ref-khashei2011}{2011}). Neural Networks mimic the
structure of the brain and consists of three layers: the input, hidden
and output layer, each of which contain nodes which are representative
of the neurons in our brain. The job of the nodes in the input layer is
to send signals along connections to the nodes in the hidden layer,
where the connections are representative of the synapses in our brain.
The nodes in the hidden layer uses an activation function and assigns a
weight to each connection according to the level of importance of
information. The value in the hidden layer is then sent along
connections to the output layer, where weights are again assigned to
each connection. The error is then computed in the output layer and is
backpropagated through the network. The weights are adjusted according
to the error that they contribute, allowing the ANN to be trained and
learn. This learning process is repeated until the network achieves the
desired output.

ANN's have had many successful applications in forecasting exchange
rates. In a study done by Fahimifard et al.
(\protect\hyperlink{ref-fahimifard2009}{2009}), the ANN was found as an
effective way to improve the forecasts of exchange rates. Superior
results were produced when compared to the ARIMA and GARCH model using
the RMSE, MSE and MAD as measure of performance. Similar results were
found by Zhang and Hu (\protect\hyperlink{ref-zhang1998}{1998}) however
the ANN failed to outperform the random walk over long forecast
horizons. Contrary to the findings above, Kuan and Liu
(\protect\hyperlink{ref-kuan1995}{1995}) attempted to investigate how
well feedforward and recurrent networks capture the non-linearity of
exchange rate data and obtained conflicting results with neural networks
having significantly better market timing ability or lower out-of-sample
Mean Square Prediction Error (MSPE) compared to the ARMA models for some
exchange rates only. Franses and Van Homelen
(\protect\hyperlink{ref-franses1998}{1998}) suggested that the
non-linear feature of exchange rates that are picked up by ANNs may
actually be due to neglected GARCH effects. They investigated the
application of ANNs to forecast exchange rates and found that the
presence of GARCH in the data may mistakenly lead one to believe that
the returns can be forecasted on the exchange rates themselves. Their
results show that there is no gain in producing out-of-sample forecasts
using ANN if the data is not truly non-linear.

Although neural networks have been broadly applied in financial
forecasting, the process of building neural networks is a complex task
and there is no consistent method to design them compared to the
traditional Box-Jenkins ARIMA model. Unlike ARIMA models, the
performance of ANNs is sensitive to many modelling factors such as the
number of input nodes and the training sample size (Zhang and Hu
\protect\hyperlink{ref-zhang1998}{1998}).

\subsection{Fuzzy Logistic Systems}\label{fuzzy-logistic-systems}

Fuzzy Logistic Systems were initially developed to solve problems
involving linguistic terms, and have been used successful in financial
forecasting (Khashei, Bijari, and Ardali
\protect\hyperlink{ref-khashei2009}{2009}). Unlike computers, the human
decision-making process includes a range of possibilities spanning `yes'
and `no'. Fuzzy logic tries to imitate this reasoning and
decision-making process by taking a discrete decision space that
consists of a large number of states and turning it into a continuous
space. This allows for more finer decisions to be provided as opposed to
discrete decisions.

Santos, Costa, and Santos Coelho
(\protect\hyperlink{ref-santos2007}{2007}) investigated how well fuzzy
systems and neural networks perform compared to the traditional ARMA and
GARCH model. He examined the forecasts of Brazilian exchange rate
returns looking at different frequencies of the series and compared
their one-step-ahead forecasts. By analysing the RMSE, U-Theil
inequality index, percentage of corrected predicted signals (CPS), and
the Pesaran-Timmermann (PT) predictive failure statistic, he found that
the Fuzzy Systems and Neural Networks achieved higher returns based on
the forecasts they produced. Similar results were found by Khashei,
Bijari, and Ardali (\protect\hyperlink{ref-khashei2009}{2009}) when he
analysed the predictive capabilities of Fuzzy Systems, Neural Networks,
the traditional ARIMA model, and a Fuzzy ARIMA model. The MAE and MSE of
the ARIMA model were higher compared to the other models.

Although Fuzzy Logistic Systems has the advantage over ARIMA models that
they can be applied to data with few observations available, it gives
acceptable rather than accurate reasoning and are more suitable for
problems which do not require high accuracy. Like ANN's there is no
systematic approach to designing Fuzzy Logistic Systems and they are
only understandable when simple.

\section{Hybrid ARIMA Models}\label{hybrid-arima-models}

Over time, many researchers began to think of ways in which to harness
the advantages of tractable linear models such as the ARIMA, and of more
complex non-linear models. By combining the ARIMA model with other
forecasting methods, the advantages of both forecasting methods are
leveraged while simultaneously improving their limitations. Some of the
hybrid models which have been commonly used are the Fuzzy ARIMA and
ANN-ARIMA model.

\subsection{Fuzzy ARIMA}\label{fuzzy-arima}

The ARIMA model produces very accurate forecasts over short time
horizons however it has the limitation of requiring more than 100
observations of historical data (Tseng et al.
\protect\hyperlink{ref-tseng2001}{2001}). In a world that is constantly
changing and with the rapid advancement of technology, access to large
amounts of historical data is difficult to obtain. On the contrary, a
fuzzy regression model requires little historical data however produces
wide prediction intervals if extreme values are present in the data. The
Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to
exploit the advantages of both models while simultaneously overcoming
their limitations. Tseng et al.
(\protect\hyperlink{ref-tseng2001}{2001}) proposed applying a Fuzzy
ARIMA model to forecast the exchange rate of Taiwan dollars to US
dollars to demonstrate the model's appropriateness and power. The Fuzzy
ARIMA not only produced forecasts that were superior to the ARIMA and
fuzzy time series models but also provided an upper and lower bound
which can be used by decision makers to determine the best and worst
possible situations.

\subsection{ANN-ARIMA}\label{ann-arima}

Although ARIMA models are powerful, they require non-stationary data to
be differenced and impose prior assumptions onto the distribution of the
data (Ince and Trafalis \protect\hyperlink{ref-ince2006}{2006}). In
contrast, machine learning techniques such as ANNs do not impose any
assumptions onto the data generating process however, being data-driven,
are sensitive to the number of input nodes used. The ANN-ARIMA model
draws on the strengths of both models and overcomes these individual
difficulties. Ince and Trafalis (\protect\hyperlink{ref-ince2006}{2006})
created an ANN-ARIMA model by using the ARIMA to determine the number
input nodes required by an ANN for three exchange rates. When the
ANN-ARIMA model was compared to the pure ARIMA, the hybrid model
outperformed the ARIMA based on the MSE. The findings of Khashei and
Bijari (\protect\hyperlink{ref-khashei2011}{2011}) study agree with
these results. They implemented an ARIMA model and used its residuals
together with past observations of the data as inputs for the ANN. Their
hybrid model had superior in-sample and out-sample forecasts compared to
the random walk, linear AR and ANN model.

@@

\section{Forecasting with Prophet}\label{forecasting-with-prophet}

@@

\subsection{Construction of Prophet}\label{construction-of-prophet}

@@

\subsection{Empirical Applications}\label{empirical-applications-1}

Prophet's Bayesian approach to forecasting allows the analyst to
incorporate their expert knowledge in the model building process and has
produced significantly improved forecasts compared to the ARIMA model.
Taylor and Letham (\protect\hyperlink{ref-taylor2017}{2017}) forecasted
the number of events on Facebook using Prophet. The time series was
impacted by holidays, had strong multi-period seasonality, and a
piecewise trend. The forecasts produced by Prophet were compared to
common automatic forecasting techniques such as exponential smoothing,
ARIMA, and the seasonal naïve model, as well as to simple models such as
the naïve model. While exponential smoothing and the seasonal naïve
model were quite robust, the ARIMA forecasts were fragile. No model
besides Prophet accounted for the dips around holidays and the upward
trend of the time series towards later observations.

@@

Prophet's ability to forecast at scale enables it to model a wide
variety of data and may be able to capture the non-linearity present in
exchange rates. In contrast, the ARIMA model fails to capture the
non-linearity inherent in exchange rate data and only produces good
forecasts over short time horizons. Prophet is also a flexible model
that has intuitive parameters which can be easily interpreted and
modified by human beings (Taylor and Letham
\protect\hyperlink{ref-taylor2017}{2017}). This may be useful to
non-experts who have domain knowledge about factors that may affect the
movement of exchange rate prices. Furthermore, unlike the ARIMA model,
Prophet can produce forecasts over irregular time intervals and allows
for missing values in the time series without the need for interpolation
(Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}). If Prophet
produces forecasts that are as good as the ARIMA model when forecasting
Bitcoin/ZAR, it can be compared to more complex forecasting methods that
are less tractable such as the hybrid models and machine learning
techniques seen earlier.

\section{Mincer-Zarnowitz Approach to Forecast
Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}

@@

\section{Conclusion}\label{conclusion}

ARIMA models have been highly popular since its introduction by Box and
Jenkins and are widely used for financial forecasting. They are easy to
implement, have a systematic procedure for model design, and produce
good forecasts over short time horizons. Although robust, they have the
limitation of requiring large amounts of historical observations and
they fail to capture the non-linearity inherent in exchange rate data.
With the evolution of computational power and the advancement of
statistics, more complex techniques such as ANN's, Fuzzy Logistic
Systems, and Hybrid ARIMA models have been proposed and compared to the
traditional ARIMA model. Although these models produce superior results
to the ARIMA model, they are less tractable and require the analyst to
understand the complex statistical intricacies involved. Prophet's
Bayesian approach to forecasting is flexible and allows for non-experts
to incorporate their domain knowledge in the model building process. If
Prophet produces forecasts that are as good as the ARIMA model when
forecasting Bitcoin/ZAR, it can be compared to these more complex
forecasting methods that are less tractable and scalable.

@@

\hypertarget{ref-cermak2017}{}
Cermak, Vavrinec. 2017. ``Can Bitcoin Become a Viable Alternative to
Fiat Currencies? An Empirical Analysis of Bitcoin's Volatility Based on
a Garch Model.''

@@

\hypertarget{ref-etuk2012}{}
Etuk, Ette Harrison, and others. 2012. ``A Seasonal Arima Model for
Daily Nigerian Naira-Us Dollar Exchange Rates.'' \emph{Asian Journal of
Empirical Research} 2 (6). Asian Economic; Social Society: 219--27.

@@

\hypertarget{ref-kuan1995}{}
Kuan, Chung-Ming, and Tung Liu. 1995. ``Forecasting Exchange Rates Using
Feedforward and Recurrent Neural Networks.'' \emph{Journal of Applied
Econometrics} 10 (4). Wiley Online Library: 347--64.

>>>>

\title{Draft}

@@

The Autoregressive Integrated Moving Average (ARIMA) model is one of the
most widely used time series models that has attracted attention in
financial market forecasting (Kashei and Bijari 2007). The ARIMA model
is known for its attractive quality of being a rich class of processes,
making it possible to find a suitable model to adequately fit the data
(Fahimifard et al. \protect\hyperlink{ref-fahimifard2009}{2009}).
Although the Random Walk model has typically been applied to foreign
exchange markets and has produced superior results, some researchers
have contended that these markets are not efficient and believe that
future prices depend on current and past events (Abu-Mostafa and Atiya
\protect\hyperlink{ref-abu1996}{1996}). This has led to the application
of the ARIMA model to exchange rate problems, where it has since gained
popularity due to its ease of implementation and tractability. Although
the ARIMA model yields good forecasts over short forecast horizons and
is easy to use, the linearity of the ARIMA model fails to adequately
capture the non-linearity inherent in exchange rate data (Zhang and Hu
\protect\hyperlink{ref-zhang1998}{1998}).

With the evolution of computational power, non-linear, soft computing
techniques have been proposed as a solution. The ARIMA model has been
used as the standard benchmark model against which these more complex
methods have been compared. Although they have produced significantly
better results than the ARIMA model, these models lack interpretability
and building them is a challenging task. The ARIMA model is more
tractable, less computationally expensive, has been used as the building
blocks for more advanced models and has provided the inspiration for
hybrid versions of the model which have been used in exchange rate
forecasting.

In December 2016, Facebook open-sourced their forecasting model Prophet.
Prophet differs to traditional time series models like ARIMA, in that it
can be utilised by non-experts who have little knowledge about the
statistical intricacies involved in the model, however have domain
knowledge about the data generating process. This can then be easily
incorporated into the model through intuitively adjustable parameters
(Taylor and Letham \protect\hyperlink{ref-taylor2017}{2017}). If Prophet
produces forecasts that are as good as the ARIMA model, it can be
compared to more complex forecasting methods that are less tractable and
flexible.

This paper broadly aims to compare the forecasts produced by the
traditional ARIMA model and Prophet through an evaluation of
Bitcoin/ZAR, using a Mincer-Zarnowitz approach of measuring forecast
accuracy. More specifically, this paper aims to:

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\item
  Compare the out-of-sample forecasts over varying forecast horizons.
\item
  Compare the forecasts produced by ARIMA's and Prophet's automatic
  selection procedures.
\item
  Compare the out-of-sample forecast performance with missing values and
  outliers present in the data.
\item
  Investigate the efficient market hypothesis (EMH) and the trading
  profit to be made with Prophet.
\end{enumerate}

Section \ref{Background} will present the key findings from the major
works in which comparisons have been made to the ARIMA model in exchange
rate forecasting, as well as the theory behind the ARIMA and Prophet
model. This will be followed by a brief explanation and motivation for
the dataset used in this paper in Section \ref{Data}. Section
\ref{Methodology} will describe the methodology used in selecting the
ARIMA and Prophet models and the Mincer-Zarnowitz approach of measuring
forecast accuracy. The results from applying the ARIMA and Prophet model
to the dataset is presented in Section \ref{Results}. Finally, Section
\ref{Discussion and Conclusions} discusses the results in a statistical
and financially meaningful context.

\section{\texorpdfstring{Background
\label{Background}}{Background }}\label{background}

\subsection{Forecasting Exchange Rates with the ARIMA
model}\label{forecasting-exchange-rates-with-the-arima-model}

@@

\[ 
    y_t = c + \sum_{i = 1}^{p} \theta_{i} y_{t-i} + \sum_{i = 1}^{q} \phi_{i} e_{t-i}+ e_t \label{eq1} \\ \notag
\] where c is a constant, \(\theta\) and \(\phi\) are parameters set
such that \ldots{} and \{\(e_t\)\} is a white noise process with zero
mean and variance \(\sigma^2\).

@@

\[ 
    (1-\sum_{i = 1}^{p} \theta_{i} B^i)(1-B)^dy_t = c + (1 + \sum_{i = 1}^{q} \phi_{i} B^i)e_t \label{eq2} \\ \notag 
\] where c is a constant and \{\(e_t\)\} is a white noise process with
zero mean and variance \(\sigma^2\) (otext.org/fpp/8/5)\ldots{}need to
get this reference hmmm

ARIMA models have been commonly used in financial forecasting and are
popular for observing stock prices and exchange rates due to its power
and statistical properties (C.-S. Lin, Chiu, and Lin
\protect\hyperlink{ref-lin2012}{2012}). They have frequently been used
as a benchmark to compare new forecasting techniques that have emerged
over time. Kamruzzaman and Sarker
(\protect\hyperlink{ref-kamruzzaman2003}{2003}) applied an ARIMA model
to forecast the exchange rate between the Australian dollar and six
other currencies, and used the forecast errors as well as the accuracy
of the direction of the forecasts to evaluate the performance of three
Neural Network models. In a similar study, Khashei, Bijari, and Ardali
(\protect\hyperlink{ref-khashei2012}{2012}) compared the predictive
capability of their proposed hybrid model consisting of an ARIMA and a
Probabilistic Neural Network (PNN), to the traditional ARIMA model, and
justified the use of the ARIMA as a benchmark by claiming that it is the
most important linear model. Their comparison was made by investigating
the forecasts between the British pound and US dollar over varying
forecast horizons. Not only has the ARIMA model been useful as a
benchmark, but it has also yielded satisfactory results when predicting
exchange rates. When Nwankwo (\protect\hyperlink{ref-nwankwo2014}{2014})
forecasted the rate between the Nairo and dollar, diagnostic testing
revealed that the ARIMA(1,0,0) model was the best fit for the data based
on Akaike's Information Criterion (AIC).

The ARIMA model is attractive as it is tractable and produces good
short-term forecasts when more than 100 observations are used (Tseng et
al. \protect\hyperlink{ref-tseng2001}{2001}). Although the ARIMA model
has the advantage of ease of implementation and flexibility, it fails to
capture the non-linearity and volatility present in exchange rate data.
Over time, the ARIMA model has evolved to cater for a wider variety of
data and to compensate for some of its shortcomings. The most popular
versions of the ARIMA model that has been implemented in exchange rate
forecasting is the Seasonal ARIMA (SARIMA) and Fractional ARIMA model
(FARIMA).

The SARIMA model was introduced to capture the periodic behaviour of
data and extends the ARIMA class by including a term for seasonal
differencing. Etuk, Wokoma, and Moffat
(\protect\hyperlink{ref-etuk2013}{2013}) modelled the Naira/CFA Franc
exchange rate which exhibited monthly seasonality using an additive
SARIMA model, to demonstrate that it can be a useful fit for exchange
rate data which displays seasonality. Their results showed that the
SARIMA model adequately described the variation in the exchange rate
series. The ARFIMA model generalises the ARIMA model in that the degree
required to make the data stationary can assume any real value, and is
no longer restricted to the integer domain. The ARFIMA model has the
ability to capture the dependence between observations that are widely
spread apart in time (Cheung \protect\hyperlink{ref-cheung1993}{1993}).
This makes the model more parsimonious since it can capture long memory
in data as well as short term dynamics (Cheung
\protect\hyperlink{ref-cheung1993}{1993}). Cheung
(\protect\hyperlink{ref-cheung1993}{1993}) fitted the ARFIMA model to
examine five exchange rates, and found that there was strong evidence of
long memory in the exchange rate time series.

Generalised autoregressive conditional heteroskedasticity (GARCH) models
were later developed in an attempt to capture the volatility in
financial markets (Anastasakis and Mort
\protect\hyperlink{ref-anastasakis2009}{2009}). Hsieh (1989) applied a
GARCH model to investigate five exchange rates and his results showed
that although the GARCH model outperformed the random walk, some
non-linear information still remained in the residuals.

\subsection{Other techniques used to forecast exchange
rates}\label{other-techniques-used-to-forecast-exchange-rates}

Over time, financial forecasting methods have moved away from linear
models like ARIMA and GARCH, to soft computing techniques. These more
complex techniques are non-linear and can fit complex time series more
easily (Castillo and Melin \protect\hyperlink{ref-castillo2002}{2002}).
Unlike the ARIMA model, soft computing techniques do not impose
structural assumptions on the model apriori (Castillo and Melin
\protect\hyperlink{ref-castillo2002}{2002}). Some of the most commonly
used artificial intelligence methods used to forecast exchange rate data
are Neural Networks and Fuzzy Logistic Systems.

Artificial Neural Networks (ANNs) have had many successful applications
in forecasting exchange rates and are more advantageous than other
non-linear forecasting methods. They are data driven, can adapt to
non-stationary environments and can approximate any continuous function
(Khashei and Bijari \protect\hyperlink{ref-khashei2011}{2011}). In a
study done by Fahimifard et al.
(\protect\hyperlink{ref-fahimifard2009}{2009}), the ANN was found as an
effective way to improve the forecasts of exchange rates. Superior
results were produced when compared to the ARIMA and GARCH model using
the root mean square error (RMSE), mean square error (MSE) and mean
absolute difference as a measure of performance. Franses and Van Homelen
(\protect\hyperlink{ref-franses1998}{1998}) suggested that the
non-linear feature of exchange rates that are picked up by ANNs may
actually be due to neglected GARCH effects. They investigated the
application of ANNs to forecast exchange rates and found that the
presence of GARCH in the data may mistakenly lead one to believe that
the returns can be forecasted on the exchange rates themselves. Their
results show that there is no gain in producing out-of-sample forecasts
using ANNs if the data is not truly non-linear. Fuzzy Logistic Systems
(FLSs) were initially developed to solve problems involving linguistic
terms, and have been successfully used in financial forecasting
(Khashei, Bijari, and Ardali \protect\hyperlink{ref-khashei2009}{2009}).
Fuzzy logic tries to imitate human reasoning and the decision-making
process and allows for finer rather discrete decisions to be provided.
Santos, Costa, and Santos Coelho
(\protect\hyperlink{ref-santos2007}{2007}) investigated how well FLSs
and ANNs perform compared to the traditional ARMA and GARCH model. He
examined the forecasts of Brazilian exchange rate returns by considering
different frequencies of the series and comparing their one step-ahead
forecasts. By analysing the RMSE, U-Theil inequality index, percentage
of corrected predicted signals (CPS), and the Pesaran-Timmermann (PT)
predictive failure statistic, he found that FLSs and ANNs achieved
higher returns based on the forecasts they produced. Similar results
were found by Khashei, Bijari, and Ardali
(\protect\hyperlink{ref-khashei2009}{2009}) when he analysed the
predictive capabilities of FLSs, ANNs, the traditional ARIMA model, and
a Fuzzy ARIMA model.

Although ANNs have been broadly applied in financial forecasting, the
process of building them is a complex task and there is no consistent
method of design compared to the traditional Box-Jenkins ARIMA model.
Unlike ARIMA models, the performance of ANNs is sensitive to many
modelling factors such as the number of input nodes included and the
size of the training sample chosen (Zhang and Hu
\protect\hyperlink{ref-zhang1998}{1998}). Like ANNs there is no
systematic approach to designing FLSs and they are only understandable
when simple. Although FLSs has the advantage over ARIMA models that they
can be applied to data with few observations available, it gives
acceptable rather than accurate results and are more suitable for
problems which do not require high accuracy.

\subsection{Hybrid ARIMA Models}\label{hybrid-arima-models}

Over time, many researchers began to think of ways in which to harness
the advantages of tractable linear models such as the ARIMA model, and
of more complex non-linear models. By combining the ARIMA model with
other forecasting methods, the advantages of both forecasting methods
are leveraged while simultaneously improving their limitations. Some of
the hybrid models which have commonly been used in exchange rate
forecasting are the Fuzzy ARIMA and ANN-ARIMA model.

The ARIMA model produces very accurate forecasts over short time
horizons however it has the limitation of requiring more than 100
observations of historical data to yield accurate results (Tseng et al.
\protect\hyperlink{ref-tseng2001}{2001}). In a world that is constantly
changing and with the rapid advancement of technology, access to large
amounts of historical data is difficult to obtain. On the contrary, a
fuzzy regression model requires little historical data however produces
wide prediction intervals if extreme values are present in the data. The
Fuzzy ARIMA model combines the ARIMA and fuzzy regression model to
exploit the advantages of both models while simultaneously overcoming
their limitations. Tseng et al.
(\protect\hyperlink{ref-tseng2001}{2001}) proposed applying a Fuzzy
ARIMA model to forecast the exchange rate of Taiwan dollars to US
dollars to demonstrate the model's appropriateness and power. The Fuzzy
ARIMA not only produced forecasts that were superior to the ARIMA and
fuzzy time series models but also provided an upper and lower bound
which can be used by decision makers to determine the best and worst
possible situations.

Although ARIMA models are powerful, they require non-stationary data to
be differenced and impose prior assumptions onto the distribution of the
data (Ince and Trafalis \protect\hyperlink{ref-ince2006}{2006}). In
contrast, machine learning techniques such as ANNs do not impose any
assumptions onto the data generating process however, being data-driven,
are sensitive to the number of input nodes used. The ANN-ARIMA model
draws on the strengths of both models and overcomes these individual
difficulties. Ince and Trafalis (\protect\hyperlink{ref-ince2006}{2006})
created an ANN-ARIMA model by using the ARIMA model to determine the
number input nodes required by an ANN for three exchange rates. When the
ANN-ARIMA model was compared to the pure ARIMA model, the hybrid model
outperformed the ARIMA based on the MSE. The findings of Khashei and
Bijari (\protect\hyperlink{ref-khashei2011}{2011}) study agree with
these results. They implemented an ARIMA model and used its residuals
together with past observations of the data as inputs for the ANN. Their
hybrid model had superior in-sample and out-sample forecasts compared to
the random walk, linear autoregressive and ANN model.

@@

\subsection{Forecasting with Prophet}\label{forecasting-with-prophet}

@@

@@

\section{\texorpdfstring{Data \label{Data}}{Data }}\label{data}

\section{\texorpdfstring{Methodology
\label{Methodology}}{Methodology }}\label{methodology}

@@

\subsection{Mincer-Zarnowitz Approach to Forecast
Evaluation}\label{mincer-zarnowitz-approach-to-forecast-evaluation}

@@

\section{\texorpdfstring{Results
\label{Results}}{Results }}\label{results}

\section{\texorpdfstring{Discussion and Conclusions
\label{Discussion and Conclusions}}{Discussion and Conclusions }}\label{discussion-and-conclusions}

@@

@@

@@
